---
title: "In Class Ex 02"

author: "Cheng Chun Chieh"

date: "20 Apr 2024"
date-modified: "last-modified"

format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
  
editor: visual
---

# 1. In Class Ex 2 - Data Visualisation

## 1.1 Loading the required packages

```{r}

pacman::p_load(tidyverse, ggdist, ggthemes, ggridges, colorspace)
```

## 1.2 Loading Exam Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# 2. Visualising Distribution

## 2.1 Typical way - Using Histogram

```{r}

ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of English scores") 
```

By using histogram - the underlying code will bin the data to create the histogram - because they are to be represented in bars.

## 2.2 Using Probability Density Plot Instead

```{r}

ggplot(data=exam_data, aes(x = ENGLISH)) +
  geom_density(color="darkblue", adjust =.65, alpha = .6, linewidth = 1) +
  theme_minimal() +
  ggtitle("Distribution of English scores") 
```

## 2.3 The alternative design

```{r}

median_eng <- median(exam_data$ENGLISH)
mean_eng <- mean(exam_data$ENGLISH)
std_eng <- sd(exam_data$ENGLISH)

ggplot(data=exam_data, aes(x = ENGLISH)) +
  geom_density(color="darkblue", adjust =.65, alpha = .6, linewidth = 1) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean_eng, sd = std_eng),
    col ="darkgrey",
    linewdith = 0.8
  ) +
  geom_vline (aes(xintercept = mean_eng), col = "red", linewidth = 0.6) + 
  annotate("text", x= mean_eng - 12, y=0.04, label= paste0("Mean ENGLISH: ", round((mean_eng),2)), color="red") + 
  annotate("text", x= median_eng + 12, y=0.04, label= paste0("Median ENGLISH: ", round((median_eng),2)), color="red") +
  geom_vline (aes(xintercept = median_eng), col = "red", linewidth = 0.6, linetype = "dashed") +
  theme_minimal() +
  ggtitle("Distribution of English scores") 
```

# 3. Using Ridgeline Plot 

[*Ridgeline plot*](https://www.data-to-viz.com/graph/ridgeline.html) (sometimes called *Joyplot*) is a data visualisation technique for revealing the distribution of a numeric value for several groups. Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.

Figure below is a ridgelines plot showing the distribution of English score by class.

-   Ridgeline plots make sense when the number of group to represent is medium to high, and thus a classic window separation would take to much space. Indeed, the fact that groups overlap each other allows to use space more efficiently. **If you have less than 5 groups, dealing with other distribution plots is probably better.**

-   It works well when there is a **clear pattern in the result**, like if there is an obvious ranking in groups. Otherwise group will tend to overlap each other, leading to a messy plot not providing any insight.

## 3.1 Plotting using ggridges method

ggridges package provides two main geom to plot gridgeline plots, they are: [`geom_ridgeline()`](https://wilkelab.org/ggridges/reference/geom_ridgeline.html) and [`geom_density_ridges()`](https://wilkelab.org/ggridges/reference/geom_density_ridges.html). The former takes height values directly to draw the ridgelines, and the latter first estimates data densities and then draws those using ridgelines.

The ridgeline plot below is plotted by using `geom_density_ridges()`.

```{r}
#| code-fold: true

ggplot(exam_data, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()

```

## 3.2 Varying Gradient in the Ridges 

We can adjust the color in the ridges using a gradient - this can be done via either [`geom_ridgeline_gradient()`](https://wilkelab.org/ggridges/reference/geom_ridgeline_gradient.html) or [`geom_density_ridges_gradient()`](https://wilkelab.org/ggridges/reference/geom_ridgeline_gradient.html). Both geoms work just like `geom_ridgeline()` and `geom_density_ridges()`, except that they allow for varying fill colors. However, they do not allow for alpha transparency in the fill. For technical reasons, we can have changing fill colors or transparency but not both.\

```{r}
#| code-fold: true

ggplot(exam_data, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Scores",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## 3.3 Mapping the Probability directly onto Colour 

Beside providing additional geom objects to support the need to plot ridgeline plot, ggridges package also provides a stat function called [`stat_density_ridges()`](https://wilkelab.org/ggridges/reference/stat_density_ridges.html) that replaces [`stat_density()`](https://ggplot2.tidyverse.org/reference/geom_density.html) of ggplot2.

Figure below is plotted by mapping the probabilities calculated by using `stat(ecdf)` which represent the empirical cumulative density function for the distribution of English score.

::: callout-caution
It is important include the argument `calc_ecdf = TRUE` in `stat_density_ridges()`.
:::

```{r}
#| code-fold: true

ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()

```

## 3.4 To Include Quantile Lines

By using [`geom_density_ridges_gradient()`](https://wilkelab.org/ggridges/reference/geom_ridgeline_gradient.html), we can colour the ridgeline plot by quantile, via the calculated `stat(quantile)` aesthetic as shown in the figure below.

```{r}
#| code-fold: true

ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_brewer(palette= "Blues", name = "Quartiles") +
  theme_ridges()
```

Instead of using number to define the quantiles, we can also specify quantiles by cut points such as 2.5% and 97.5% tails to colour the ridgeline plot as shown in the figure below.

```{r}
ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("lightblue", "lightgrey", "darkblue"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]"),
    ) +
  theme_ridges()
```

# 4. Using Raincloud Plot

Raincloud Plot is a data visualisation techniques that produces a half-density to a distribution plot. It gets the name because the density plot is in the shape of a “raincloud”. The raincloud (half-density) plot enhances the traditional box-plot by highlighting multiple modalities (an indicator that groups may exist). The boxplot does not show where densities are clustered, but the raincloud plot does!

In this section, you will learn how to create a raincloud plot to visualise the distribution of English score by race. It will be created by using functions provided by **ggdist** and ggplot2 packages.

## 4.1 Plotting a Half Eye Graph

First, we will plot a Half-Eye graph by using [`stat_halfeye()`](https://mjskay.github.io/ggdist/reference/stat_halfeye.html) of **ggdist** package.

This produces a Half Eye visualization, which is contains a half-density and a slab-interval.

```{r}
#| code-fold: 

ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

::: callout-tip
We remove the slab interval by setting .width = 0 and point_colour = NA.
:::

## 4.2 Adding in the boxplot

Next, we will add the second geometry layer using [`geom_boxplot()`](https://r4va.netlify.app/chap09) of ggplot2. This produces a narrow boxplot. We reduce the width and adjust the opacity.

```{r}
#|code-fold: true

ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

## 4.3 Adding the dot plots

Next, we will add the third geometry layer using [`stat_dots()`](https://mjskay.github.io/ggdist/reference/stat_dots.html) of ggdist package. This produces a half-dotplot, which is similar to a histogram that indicates the number of samples (number of dots) in each bin. We select side = “left” to indicate we want it on the left-hand side.

```{r}
#| code-fold: true

ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

## 4.4 Finishing Touches 

Lastly, [`coord_flip()`](https://ggplot2.tidyverse.org/reference/coord_flip.html) of ggplot2 package will be used to flip the raincloud chart horizontally to give it the raincloud appearance. At the same time, `theme_economist()` of ggthemes package is used to give the raincloud chart a professional publishing standard look.

```{r}
#| code-fold: true

ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

# References

-   [Introducing Ridgeline Plots (formerly Joyplots)](https://blog.revolutionanalytics.com/2017/07/joyplots.html)

-   Claus O. Wilke [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/) especially Chapter [6](https://clauswilke.com/dataviz/visualizing-amounts.html), [7](https://clauswilke.com/dataviz/histograms-density-plots.html), [8](https://clauswilke.com/dataviz/ecdf-qq.html), [9](https://clauswilke.com/dataviz/boxplots-violins.html) and [10](https://clauswilke.com/dataviz/visualizing-proportions.html).

-   Allen M, Poggiali D, Whitaker K et al. [“Raincloud plots: a multi-platform tool for robust data. visualization”](https://wellcomeopenresearch.org/articles/4-63) \[version 2; peer review: 2 approved\]. Welcome Open Res 2021, pp. 4:63.

-   [Dots + interval stats and geoms](https://mjskay.github.io/ggdist/articles/dotsinterval.html)
